---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# r2mlm

<!-- badges: start -->
[![CRAN_Status_Badge](https://img.shields.io/cran/v/r2mlm)](https://cran.r-project.org/package=r2mlm)
<!-- badges: end -->

The r2mlm package contains functions that compute a framework of total and level-specific R-squared measures for multilevel models, with accompanying plots; these plots allow interpreting and visualizing all of the measures together as an integrated set. The framework of R-squared measures subsumes and analytically relates 10 previously-developed measures as special cases of 5 measures from this framework, as well as provides several new measures. The framework is presented in Rights & Sterba (2019) for evaluating a single fitted multilevel model. The implementation of this framework of measures for comparing multilevel models using R-squared differences is described in Rights & Sterba (2020). The functions in this package allow users to input either model objects obtained from [lme4](https://cran.r-project.org/web/packages/lme4/index.html) or [nlme](https://cran.r-project.org/web/packages/nlme/index.html), or to manually input model parameter estimates.

## Installation

You can install the released version of r2mlm from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("r2mlm")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mkshaw/r2mlm")
```
## Example

Suppose you have a dataset consisting of teachers nested within schools. A researcher could specify a multilevel model with teacher job satisfaction as the outcome, which is predicted by the level-1 predictor teacher salary (here school-mean-centered) and the level-2 predictor student-teacher ratio. Suppose the multilevel model included a random intercept as well as a random slope of teacher salary, and included normally-distributed, homoscedastic level-1 residuals. The researcher could then obtain the following output (see r2mlm function for further details). 

```{r example}
library(r2mlm)

# Generate the model, in this case with lme4:
model <- lmer(satisfaction ~ 1 + salary_c + s_t_ratio + (1 + salary_c | schoolID), data = teachsat, REML = TRUE)

# Adjust plot margins
par(mar = c(7.1, 6.1, 6.1, 4.1))

# Generate R-squared measures for that model:
r2mlm(model)
```

## Package Functions

There are two main functions currently available in r2mlm:

1. `r2mlm()`, for computing variance explained for a single multilevel model.
2. `r2mlm_comp()`, for comparing variance explained between two different multilevel models.

In some cases, you might run a multilevel model that will not converge in `nlme` or `lme4`, but will converge in another software package (e.g., HLM, Mplus). If you run your model and obtain the associated output, you can manually generate R-squared measures using `r2mlm_manual()` or `r2mlm_comp_manual()`. For manual entry details, see the help pages:

``` r
?r2mm_manual()
?r2mlm_comp_manual()
```

## Framework Assumptions

This framework of variance explained assumes the following:

1. Two-level multilevel linear models (3 or more levels not supported at this time);
2. Normal outcome variable;
3. Homoscedastic residual variances at both level 1 and level 2.

## References

Rights, J. D., & Sterba, S. K. (2019). Quantifying explained variance in multilevel models: An integrative framework for defining R-squared measures. *Psychological Methods*, *24*(3), 309–338. https://doi.org/10.1037/met0000184 

Rights, J. D., & Sterba, S. K. (2020). New recommendations on the use of R-squared differences in multilevel model comparisons. *Multivariate Behavioral Research*. https://doi.org/10.1080/00273171.2019.1660605


