---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# r2mlm

<!-- badges: start -->
[![CRAN_Status_Badge](https://img.shields.io/cran/v/r2mlm)](https://cran.r-project.org/package=r2mlm)
<!-- badges: end -->

r2mlm generates R-squared measures for multilevel models, based on the framework described in [Rights & Sterba (2019)](https://psycnet.apa.org/record/2018-32178-001). You provide a multilevel model generated using [lme4](https://cran.r-project.org/web/packages/lme4/index.html) or [nlme](https://cran.r-project.org/web/packages/nlme/index.html), and r2mlm outputs both total- and level-specific variance explained. 

## Installation

You can install the released version of r2mlm from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("r2mlm")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mkshaw/r2mlm")
```
## Example

As an example, say you have a dataset of teachers clustered within schools. Your model has teacher job satisfaction as the outcome, predicted by teacher salary (within-school, level 1 variable) and student-teacher ratio (between-school, level 2 variable).

```{r example}
library(r2mlm)

# Generate the model, in this case with lme4:
model <- lmer(satisfaction ~ 1 + salary_c + s_t_ratio + (1 | schoolID), data = teachsat, REML = TRUE)

# Generate R-squared measures for that model:
r2mlm(model)
```

## Package Functions

There are two main functions currently available in r2mlm:

1. `r2mlm()`, for computing variance explained for a single model.
2. `r2mlm_comp()`, for comparing variance explained between two different models.

In some cases, you might run a model that will not converge in `nlme` or `lme4`, but will converge in another software (e.g., HLM, Mplus). If you run your model and have the associated output, you can manually generate R-squared measures using `r2mlm_manual()` or `r2mlm_comp_manual()`. For manual entry details, see the help pages:

``` r
?r2mm_manual()
?r2mlm_comp_manual()
```

## Framework Assumptions

This framework of variance explained assumes the following:

1. Two-level multilevel linear models (3 or more levels not supported at this time);
2. Normal outcome variable;
3. Homoscedastic residual variances at both level 1 and level 2;
4. Level-1 predictors are centered within cluster.

## Getting Help

